Equalizer\\Programming and User Guide

### 2010-02-22 Developer/ProgrammingGuide/paper.tex
Programming Guide: add slave object commit section
[ ] May break build
[ ] May break existing applications
[x] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[ ] Documentation

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications based on
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.7.3 for Equalizer 0.9.1\\[\medskipamount]
%  \today
}

@@

Written by Stefan Eilemann.\\
Engineering contributions by Maxim Makhinya, Jonas B\"osch, Christian
Marten, Sarah Amsellem, Patrick Bouchaud and Philippe Robert.

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.7.0   & Sep 09, 2009 & Updates based on hardcopy print\\
    1.7.1   & Oct 20, 2009 & Updates for new eq::GLWindow\\
    1.7.2   & Nov 18, 2009 & Add \sref{sSubpixel} and \sref{sFSAA} for Subpixel compounds\\
    1.7.3   & Feb 02, 2010 & Revision for Equalizer 0.9.1\\
    \hline
  \end{tabularx}}
\clearpage

@@

Equalizer provides distributed objects which facilitate the
implementation of data distribution in a cluster
environment. Distributed objects are created by subclassing from
\textsf{eq::Object} or \textsf{eq::net::Object}. The application
programmer implements serialization and deserialization of the
distributed data.

Objects are dynamically assigned to a session, which assigns
session-unique identifiers to objects and manages the network
communication between different instances of the same distributed
object.

Distributed objects can be static (immutable) or dynamic. Dynamic
objects are versioned.

The \textsf{eqPly} example uses static distributed objects to provide
initial data and the model to all rendering nodes, as well as a
versioned object to provide frame-specific data such as the camera
position to the rendering methods.

\subsubsection{Common Usage}

\begin{wrapfigure}{r}{.382\textwidth}
%\begin{figure}[h!t]
  \subfigure[]{\label{fObjectDist1}
    \includegraphics[width=.382\textwidth]{images/umlObjectSubclass}}\hfil
  \subfigure[]{\label{fObjectDist2}
    \includegraphics[width=.382\textwidth]{images/umlObjectProxy}}\hfil
  \subfigure[]{\label{fObjectDist3}
    \includegraphics[width=.382\textwidth]{images/umlObjectMI}}%
  {\caption{\label{fObjectDist}Object Distribution using Subclassing,
      Proxies or Multiple Inheritance}}
\end{wrapfigure}
Distributed objects are addressed using session-unique identifiers,
because pointers to other objects cannot be distributed directly; they
have no meaning on remote nodes. The session used to register
distributed objects is normally the \textsf{eq::Config}, which is
sub-classed from \textsf{eq::net::Session}.

The entry point for shared data on a render client is the identifier
passed by the application to \textsf{Config::init}. This identifier
typically contains the identifier of a static distributed object, and is
passed by Equalizer to all \textsf{configInit} task methods. Normally
this initial object is mapped by the render clients in
\textsf{Node::configInit}. It typically contains identifiers of other
shared data objects.

There are multiple ways of implementing data distribution for an
existing C++ class hierarchy:
\begin{description}
\item[Subclassing] The classes to be distributed inherit from
  \textsf{eq::net::Object} and implement the serialization methods. This
  approach is recommended if the source code of existing classes can be
  modified. It is used for \textsf{eqPly::InitData} and
  \textsf{eqPly::FrameData}. (\fig{fObjectDist1})
\item[Proxies] For each object to be distributed, a proxy object is
  created which manages data distribution for its associated
  object. This requires the application to track changes on the object
  separately from the object itself. The model data distribution of
  eqPly is using this pattern. (\fig{fObjectDist2})
\item[Multiple Inheritance] A new class inheriting from the class to be
  distributed and from \textsf{eq::net::Object} implements the data
  distribution. This requires the application to instantiate a different
  type of object instead of the existing object, and to create wrapper
  methods in the superclass calling the original method and setting the
  appropriate dirty flags. This pattern is not used in
  eqPly. (\fig{fObjectDist3})
\end{description}

\subsubsection{Change Handling}

Equalizer determines the way changes are to be handled by calling
\textsf{Object::get\-Change\-Type} during the registration of the master
version of a distributed object. The change type determines the memory
footprint and the contract for the serialization methods. The following
change types are possible:

\begin{description}
  \item[STATIC] The object is not versioned. The instance data is
    serialized whenever a new slave instance is mapped. No additional
    data is stored.
  \item[INSTANCE] The object is versioned, and the instance and delta
    data is identical, that is, only instance data is
    serialized. Previous instance data is saved to be able to map old
    versions.
  \item[DELTA] The object is versioned, and the delta data is typically
    smaller than the instance data. Both the delta and instance data are
    serialized and saved to map old versions.
  \item[UNBUFFERED] The object is versioned. No data is stored, and no
    previous versions can be mapped. The instance data is serialized
    whenever a new slave instance is mapped. The application can choose
    to use a different, more optimal implementation to pack deltas.
\end{description}

@@

\begin{itemize}
\item Only the master instance of the object is writable, that is,
  \textsf{eq::net::Object::com\-mit} can be called only on the master
  instance to generate a new version.
\item Slave instance versions can only be advanced, that is,
  \textsf{eq::net::Object::sync( version )} with a version smaller than
  the current version will fail.
\item Newly mapped slave instances are mapped to the oldest available
  version by default, or to the version specified when calling
  \textsf{Session::mapObject}.
\end{itemize}

@@

\begin{wrapfigure}{r}{.618\textwidth}
%\begin{figure}[ht!]\center
  \includegraphics[width=.618\textwidth]{images/objects.pdf}
  {\caption{\label{fObjects}Scene Graph Distribution}}
\end{wrapfigure}
Applications distributing a dynamic scene graph use the frame data
instead of the init data as the entry point to their scene graph data
structure. \fig{fObjects} shows one possible implementation, where the
identifier and version of the scene graph root are transported using the
frame data. The scene graph root then serializes and de-serializes his
immediate children by transferring their identifier and current version,
similar to the static distribution done by \textsf{eqPly}.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2ex}\includegraphics[width=.618\textwidth]{images/osWindow.pdf}
  {\caption{\label{fOSWindow}eq::OSWindow UML Class Hierarchy}}\vspace{-4ex}
\end{wrapfigure}
The particulars of creating a window and OpenGL context depend on the
window system used. One can either use the implementation provided by
the operating system, e.g., AGL, WGL or glX, or some higher-level
toolkit, e.g., Qt.

@@

{\footnotesize\begin{lstlisting}
struct ConfigEvent : public eq::ConfigEvent
{
public:
    enum Type
    {
        READBACK = eq::ConfigEvent::USER,
        ASSEMBLE
    };

@@

Changing view frustum parameters is typically done for non-fullscreen
application windows and multi-view layouts, where the rendering is not
meant to be viewed in real-world size in an immersive environment. A
typical use case is changing the field-of-view of the rendering.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/viewUML.pdf}
  {\caption{\label{fViewUML}UML Hierarchy of \textsf{eqPly::View}}}
\end{wrapfigure}
A view is derived from \textsf{eq::Frustum} (\fig{fViewUML}), and the
application process can set the wall or projection parameters at
runtime. For a description of wall and projection parameters please refer
to \sref{sFrustum}. The new data will be effective for the next frame.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\label{fDirectSend}Direct Send Compositing}}
\end{wrapfigure}
The main idea behind direct send is to parallelize the costly
recomposition for database (sort-last) decomposition. With each
additional source channel, the amount of pixel data to be composited
grows linearly. When using the simple approach of compositing all frames
on the destination channel, this channel quickly becomes the bottleneck
in the system. Direct send distributes this workload evenly across all
source channels, and thereby keeps the compositing work per channel
constant.

In direct send compositing, each rendering channel is also responsible
for the sort-last composition of one screen-space tile. He receives
the framebuffer pixels for his tile from all the other channels. The
size of one tile decreases linearly with the number of source channels,
which keeps the total amount of pixel data per channel constant.

@@

The offset of input and output frames characterizes the position of the
frame data relative to the framebuffer, that is, the \textbf{window's}
lower-left corner. For output frames this is the position of the channel
relative to the window.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\label{fSlices}Volume Rendering using 3D Textures}}
\end{wrapfigure}
The \textsf{eVolve} example is a scalable volume renderer. It uses 3D
texture-based volume rendering, where the 3D texture is intersected by
view-aligned slices. The slices are rendered back-to-front and blended
to produce the final image, as shown in \fig{fSlices}\footnote{Volume
  Data Set courtesy of: SFB-382 of the German Research Council (DFG)}.

@@

\subsection{Session}

@@

\subsection{\label{sNetObject2}Object}

Distributed objects provide powerful, object-oriented data
distribution for C++ objects. Their functionality and an example use
case for parallel rendering has been described in \sref{sNetObject}.

The master instance of an \textsf{eq::net::Object} is registered with a
session, which assigns a session-unique identifier to the object. Using
this identifier, other nodes can map their slave instance of the object
to their local proxy of the session.

@@

Multiple instances of the same object (identifier) can be mapped on the
same node and session simultaneously. Each object instance has a
node-unique instance identifier. When the instance ID is set to
\textsf{EQ\_ID\_ANY} (the default value), a command is dispatch to all
object instance registered to the session with the correct object
ID. Multiple object instance per node are used when different threads on
the node need to process a different version of the same object.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/netObject.pdf}
  {\caption{\label{fNetObject}Communication between two Objects}}
\end{wrapfigure}
Distributed objects provide data synchronization functionality, as
described in \sref{sNetObject}. Additionally, applications can send
command packets by creating custom object packets with a command greater
than \textsf{CMD\_OBJECT\_CUSTOM}.

>>>>

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications based on
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.7.4 for Equalizer 0.9.1+svn\\[\medskipamount]
%  \today
}

@@

Written by Stefan Eilemann.\\
Engineering contributions by Maxim Makhinya, Jonas B\"osch, Christian
Marten, Sarah Amsellem, Patrick Bouchaud, Philippe Robert and Robert Hauck.

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
\clearpage
\listoffigures \vfill
\section*{Revision History}
{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.6     & Aug 07, 2009 & Revision for Equalizer 0.9\\
    1.7.0   & Sep 09, 2009 & Updates based on hardcopy print\\
    1.7.1   & Oct 20, 2009 & Updates for new eq::GLWindow\\
    1.7.2   & Nov 18, 2009 & Add \sref{sSubpixel} and \sref{sFSAA} for Subpixel compounds\\
    1.7.3   & Feb 02, 2010 & Revision for Equalizer 0.9.1\\
    1.7.4   & Feb 22, 2010 & Add \sref{sSlaveCommit}\\
    \hline
  \end{tabularx}}
\clearpage

@@

Equalizer provides distributed objects which facilitate the
implementation of data distribution in a cluster
environment. Distributed objects are created by subclassing from
\textsf{eq::Object} or \textsf{eq::net::Object}. The application
programmer implements serialization and deserialization of the
distributed data. \sref{sNetObject2} covers distributed objects in detail.

Distributed objects can be static (immutable) or dynamic. Dynamic objects are
versioned. The \textsf{eqPly} example uses static distributed objects to provide
initial data and the model to all rendering nodes, as well as a versioned object
to provide frame-specific data such as the camera position to the rendering
methods.

@@

\begin{itemize}
\item Only the master instance of the object generates new versions, that is,
  \textsf{eq::net::Object::com\-mit} should only be called only on the master
  instance. It is possible to commit on slave instances, but special care has to
  be taken to handle possible conflicts. \sref{sSlaveCommit} covers slave object
  commits in detail.
\item Slave instance versions can only be advanced, that is,
  \textsf{eq::net::Object::sync( version )} with a version smaller than
  the current version will fail.
\item Newly mapped slave instances are mapped to the oldest available
  version by default, or to the version specified when calling
  \textsf{Session::mapObject}.
\end{itemize}

@@

%\begin{wrapfigure}{r}{.618\textwidth}
\begin{figure}[ht!]\center
  \includegraphics[width=.618\textwidth]{images/objects.pdf}
  {\caption{\label{fObjects}Scene Graph Distribution}}
\end{figure}
Applications distributing a dynamic scene graph use the frame data
instead of the init data as the entry point to their scene graph data
structure. \fig{fObjects} shows one possible implementation, where the
identifier and version of the scene graph root are transported using the
frame data. The scene graph root then serializes and de-serializes his
immediate children by transferring their identifier and current version,
similar to the static distribution done by \textsf{eqPly}.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2ex}\includegraphics[width=.618\textwidth]{images/osWindow.pdf}
  {\caption{\label{fOSWindow}eq::OSWindow UML Class Hierarchy}}
\end{wrapfigure}
The particulars of creating a window and OpenGL context depend on the
window system used. One can either use the implementation provided by
the operating system, e.g., AGL, WGL or glX, or some higher-level
toolkit, e.g., Qt.

@@

{\footnotesize\begin{lstlisting}
struct ConfigEvent : public eq::ConfigEvent
{
public:
    enum Type
    {
        READBACK = eq::Event::USER,
        ASSEMBLE
    };

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/viewUML.pdf}
  {\caption{\label{fViewUML}UML Hierarchy of \textsf{eqPly::View}}}
\end{wrapfigure}
Changing view frustum parameters is typically done for non-fullscreen
application windows and multi-view layouts, where the rendering is not
meant to be viewed in real-world size in an immersive environment. A
typical use case is changing the field-of-view of the rendering.

A view is derived from \textsf{eq::Frustum} (\fig{fViewUML}), and the
application process can set the wall or projection parameters at
runtime. For a description of wall and projection parameters please refer
to \sref{sFrustum}. The new data will be effective for the next frame.

@@

The main idea behind direct send is to parallelize the costly
recomposition for database (sort-last) decomposition. With each
additional source channel, the amount of pixel data to be composited
grows linearly. When using the simple approach of compositing all frames
on the destination channel, this channel quickly becomes the bottleneck
in the system. Direct send distributes this workload evenly across all
source channels, and thereby keeps the compositing work per channel
constant.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/directSend.pdf}
  {\caption{\label{fDirectSend}Direct Send Compositing}}
\end{wrapfigure}
In direct send compositing, each rendering channel is also responsible
for the sort-last composition of one screen-space tile. He receives
the framebuffer pixels for his tile from all the other channels. The
size of one tile decreases linearly with the number of source channels,
which keeps the total amount of pixel data per channel constant.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/assembly.pdf}
  {\caption{\label{fAssembly}Hierarchy of Assembly Classes}}
\end{wrapfigure}
The offset of input and output frames characterizes the position of the
frame data relative to the framebuffer, that is, the \textbf{window's}
lower-left corner. For output frames this is the position of the channel
relative to the window.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\label{fSlices}Volume Rendering using 3D Textures}}\vspace{-2ex}
\end{wrapfigure}
The \textsf{eVolve} example is a scalable volume renderer. It uses 3D
texture-based volume rendering, where the 3D texture is intersected by
view-aligned slices. The slices are rendered back-to-front and blended
to produce the final image, as shown in \fig{fSlices}\footnote{Volume
  Data Set courtesy of: SFB-382 of the German Research Council (DFG)}.

@@

\subsection{Sessions}

@@

\subsection{\label{sNetObject2}Objects}

Distributed objects provide powerful, object-oriented data distribution for C++
objects. They facilitate the implementation of data distribution in a cluster
environment. Their functionality and an example use case for parallel rendering
has been described in \sref{sNetObject}.

Distributed objects are created by subclassing from \textsf{eq::Object} or
\textsf{eq::net::Object}. The application programmer implements serialization
and deserialization of the distributed data. Objects are dynamically assigned to
a session, which assigns session-unique identifiers to objects and manages the
network communication between different instances of the same distributed
object.

Distributed objects can be static (immutable) or dynamic. Dynamic objects are
versioned. New versions are commited from the master instance, which sends
the delta between the previous and current version to all mapped slave
objects. The slave objects sync the queued deltas when they need a version.

\subsubsection{Common Usage for Parallel Rendering}

Distributed objects are addressed using session-unique identifiers,
because pointers to other objects cannot be distributed directly; they
have no meaning on remote nodes. The session used to register
distributed objects is normally the \textsf{eq::Config}, which is
sub-classed from \textsf{eq::net::Session}.

The entry point for shared data on a render client is the identifier
passed by the application to \textsf{Config::init}. This identifier
typically contains the identifier of a static distributed object, and is
passed by Equalizer to all \textsf{configInit} task methods. Normally
this initial object is mapped by the render clients in
\textsf{Node::configInit}. It normally contains identifiers of other
shared data objects.

\begin{wrapfigure}{r}{.382\textwidth}
%\begin{figure}[h!t]
  \subfigure[]{\label{fObjectDist1}
    \includegraphics[width=.382\textwidth]{images/umlObjectSubclass}}\hfil
  \subfigure[]{\label{fObjectDist2}
    \includegraphics[width=.382\textwidth]{images/umlObjectProxy}}\hfil
  \subfigure[]{\label{fObjectDist3}
    \includegraphics[width=.382\textwidth]{images/umlObjectMI}}%
  {\caption{\label{fObjectDist}Object Distribution using Subclassing,
      Proxies or Multiple Inheritance}}
\end{wrapfigure}
There are multiple ways of implementing data distribution for an
existing C++ class hierarchy, such as a scene graph:
\begin{description}
\item[Subclassing] The classes to be distributed inherit from
  \textsf{eq::net::Object} and implement the serialization methods. This
  approach is recommended if the source code of existing classes can be
  modified. It is used for \textsf{eqPly::InitData} and
  \textsf{eqPly::FrameData}. (\fig{fObjectDist1})
\item[Proxies] For each object to be distributed, a proxy object is
  created which manages data distribution for its associated
  object. This requires the application to track changes on the object
  separately from the object itself. The model data distribution of
  eqPly is using this pattern. (\fig{fObjectDist2})
\item[Multiple Inheritance] A new class inheriting from the class to be
  distributed and from \textsf{eq::net::Object} implements the data
  distribution. This requires the application to instantiate a different
  type of object instead of the existing object, and to create wrapper
  methods in the superclass calling the original method and setting the
  appropriate dirty flags. This pattern is not used in
  eqPly. (\fig{fObjectDist3})
\end{description}

\subsubsection{Change Handling and Serialization}

Equalizer determines the way changes are to be handled by calling
\textsf{Object::get\-Change\-Type} during the registration of the master version
of a distributed object. The change type determines the memory footprint and the
contract for the serialization methods. The deserialization always happens in
the application thread causing the deserialization. The following change types
are possible:

\begin{description}
  \item[STATIC] The object is not versioned. The instance data is serialized
    whenever a new slave instance is mapped. The serialization happens in the
    command thread on the master node, i.e., in parallel to the application
    threads. Since the object's distributed data is supposed to be static, this
    should not create any race conditions. No additional data is stored.
  \item[INSTANCE] The object is versioned, and the instance and delta data is
    identical, that is, only instance data is serialized. The serialization
    always happens before \textsf{Session::registerObject} or
    \textsf{Object::commit} returns. Previous instance data is saved to be able
    to map old versions.
  \item[DELTA] The object is versioned, and the delta data is typically smaller
    than the instance data. Both the delta and instance data are serialized
    before \textsf{Object::commit} returns. Previous instance data is saved to
    be able to map old versions.
  \item[UNBUFFERED] The object is versioned. No data is stored, and no previous
    versions can be mapped. The instance data is serialized from the command
    thread whenever a new slave instance is mapped, i.e., in parallel to
    application threads. The application has to ensure that this does not create
    any thread conflicts. The delta data is serialized before
    \textsf{Object::commit} returns. The application may choose to use a
    different, more optimal implementation to pack deltas by using a different
    implementation for \textsf{getInstanceData} and \textsf{pack}.
\end{description}

\subsubsection{\label{sSlaveCommit}Slave Object Commit}

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-2ex}\includegraphics[width=.618\textwidth]{images/slaveWrites.pdf}
  {\caption{\label{fSlaveCommit}Slave Commit Communication Sequence}}
\end{wrapfigure}
Versioned slave objects may also commit data to the master instance. However,
this requires the implementation to take special care to resolve conflicts
during deserialization. The algorithm and constraints for slave object commits
are explained in this section.

When using \textsf{commit} on slave object instances, multiple modifications on
the same database may happen simultaneously. In order to avoid excessive
synchronization in the network layer, Equalizer a light-weight approach which
requires the application to avoid or handle conflicting changes on the same
object.

When a slave object instance commits data, \textsf{pack} is called on the
object. The data serialized by \textsf{pack} is send to the master instance of
the object. On the master instance, \textsf{notifyNewVersion} is called from the
command thread to signal the data availability. No new version is generated on
the slave instance.

The master instance can \textsf{sync} delta data received from slaves. The data
is synchronized in the order it has been received from the slaves. If a specific
order is needed, it is the application's responsibity to serialize the slave
commits accordingly. When using \textsf{VERSION\_NEXT} as the version parameter
to \textsf{sync}, one slave commit is unpacked. This call may block if no data
is pending. When using \textsf{VERSION\_HEAD} all pending versions are
unpacked. This call will never block.

Synchronizing slave data on a master instance does not create a new version. The
application has to explicitly call \textsf{commit} to pack the changed data,
which will create a new version send to all slaves.

This approach requires the application to handle or avoid, if possible, the
following situations:
\begin{itemize}
\item A slave instance will always receive its own changes when the master
  commits the slave change.
\item A slave instance commits data which does not correspond to the last
  version of the master instance.
\item Two different slave instances commit changes simultaneously which may
  result in conflicts.
\item A master instance applies data to an already changed database.
\item The master instance commits a new version while a slave instance is
  modifying its instance. 
\end{itemize}

\fig{fSlaveCommit} depicts one possible communication sequence. In general, it
is best to adhere to the following rules:
\begin{itemize}
\item Handle the case that a slave instance will receive its own changes again.
\item Write your application code so that only one object instance is modified
  at any given time. 
\item When multiple slave instances have to modify the same object at the same
  time, try to modify and serialize disjunct data of the object.
\item Before modifying a slave instance, \textsf{sync} it to the head version.
\item Write serialization code defensively, e.g.:
  \begin{itemize}
  \item When adding a new child to an object, check that it has not been added
    already.
    \item When removing a child, transmit the child's identifier instead of its
      index for removal.
    \item Modify a variable by serializing its value (instead of an operation on
      it) so that the last change will win.
  \end{itemize}
\end{itemize}

\subsubsection{Communication between Object Instances}

@@

Multiple instances of the same object (identifier) can be mapped on the
same node and session simultaneously. Each object instance has a
node-unique instance identifier. When the instance ID is set to
\textsf{EQ\_ID\_ANY} (the default value), a command is dispatch to all
object instances registered to the session with the correct object
ID. Multiple object instance per node are used when different threads on
the node need to process a different version of the same object.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/netObject.pdf}
  {\caption{\label{fNetObject}Communication between two Objects}}
\end{wrapfigure}
Distributed objects provide data synchronization functionality, as described
above. Additionally, applications can send command packets by creating custom
object packets with a command greater than \textsf{CMD\_OBJECT\_CUSTOM}.
